Parallel training by Pipeline Execution In AIRCoP

### 2021-12-12 report/paper.tex
Second proof read and Minor Grammar fix

\end{figure} This overhead may be negligible for a GPU Cluster with high-bandwidth interconnections but will cause a significant performance drop for large distributed systems consisting of several nodes of GPU clusters with low bandwidth connection between them Figure \ref{fig:networktopo}. Because high bandwidth inter-GPU connection techniques like NV-Link can only scale to a limited number of devices, large training servers are often deployed across multiple nodes. It will be ideal to divide the model into several modules, assign each module to a cluster, use Spatial partition inside a GPU cluster, and adopt pipeline parallelism for modules across different clusters.\par

@@

 AIRCoP \cite{aircopfeiwang} is the first framework that supports all the parallel parallelisms above, therefore allows users to fully explore different parallel settings. AIRCoP implemented spatial partition before and our work implemented Pipeline parallelism for AIRCoP, thus make AIRCoP the first unifying parallel training framework.\par

@@

 AIRCoP developed a formal syntax to represent tensor computation graphs. The syntax mainly has three components: Tensor and corresponding TensorType to represent input tensor and weight, Operation, and corresponding Operation Type to represent different Machine learning operations like Conv/MatMul/Relu and their input/output Type, Module to represent different modules. Operations and Modules are connected through let-bindings. Furthermore, we use annotation of the formal syntax to represent different parallelism strategies formally. We developed annotation systems for both Tensor, Operation, and Modules. Tensor and Operation annotation represents Spatial Partition by indicating the dimension to partition. Module annotation indicates how many micro-batches the mini-batch is split into. Users only need to write the sequential forward computation of the Machine Learning model, and the correct final training loop will be automatically computed through a series of transformations.

@@

 Simply splitting modules will cause large amount of stall time (or bubble time) due to complex dependencies between the sequential modules. As shown in Figure \ref{fig:multiplemodule}, if we naively spilt the model into 4 modules $m_0,m_1,m_2,m_3$. There will be sequential dependencies between the forward pass $F_0, F_1, F_2, F_3$ of modules $m_0,m_1,m_2,m_3$, and reverse dependencies between backward pass $B_3, B_2, B_1, B_0$. The dependencies between the forward pass and backward pass will force strict ordering between modules, introduce synchronizations between module boundaries, and cause significant idle time on 4 devices. \par

@@

 LMS (Lightweight Modular Staging) is a library-based staging compiler that distinguishes stages solely based on Type information. LMS is beneficial for developing DSL through overloading. It has a typed frontend making the code generation safe. LMS offers lifted (next-stage) versions of regular operators, and standard control-flow statements make symbolic code generation easy and convenient. AIRCoP is simply a tensor computation DSL embedded inside LMS framework.\par

@@

The Collection Phase collects the forward operations, the weight nodes, and the corresponding backward operation of each node. The backward operation is collected in reverse order to maintain the correct semantic of differentiation. Because the backward pass should be generated after the forward pass, we collect backward operation in the form of closures. We implement the Collection Phase by defining a function aircopCollect for each IR Node to collect the corresponding information. After the Collection Phase, we have the necessary information to perform automatic differentiation (AD). We also need to collect lifted tensors to generate the correct training loop. As we mentioned before, the forward pass and backward pass of micro-batches are iterated in different loops to achieve pipeline parallelism. Therefore, the tensors allocated in the forward pass and used in the backward pass will cause an error because forward and backward passes are now in different scopes. These tensors are called lifted tensors. We have to lift them into the outer scope and maintain multiple versions of them as an array of tensors. Micro-batches' forward and backward passes will use different versions of the lifted tensors. After collecting the lifted tensors, the Collection Phase is finished.\par

@@

The Generation Phase generates the correct training loop based on the collected information. It will add paired send/recv operations at the boundaries of different modules to maintain the correct data dependencies. Inside each module, it will first generate the top-level training loop for mini-batches. Then it will generate the forward training loop and backward training loop for micro-batches based on Module Annotations and use the collected backward operations to realize automatic differentiation as we mentioned in Section \ref{formalpipe}. Inside both the forward and backward pass, the read and write to the lifted tensors will be transformed into read/write to the versioned tensor array. The weight gradient will be accumulated in the backward pass and optimized at the end of the entire mini-batch.\par

@@

 This section shows the evaluation benchmarks of various forms of Pipeline parallelism by running our generated CUDA code on 4 GeForce TITAN XP GPUs or server cuda.cs.purdue.edu. The server is a single GPU cluster with a high-speed inter-GPU connection. The ideal network topology for Pipeline parallelism is an extensive training system consisting of several Nodes, but we currently could not run the evaluation under that scenario due to limited hardware resources and time. The purpose of this benchmark is to validate our graph transformation and gain a deeper insight into the choice of different parallelism strategies.\par

@@

 The running time of 5 settings is shown in figure\ref{fig:runtime}, the corresponding speedup is shown in figure\ref{fig:speedup}. We can notice that simply dividing a model into 2 modules will not cause any speed up due to the low device usage caused by the cross-module dependencies. The device will idle while waiting for data to arrive. We can notice that using Pipeline parallelism with 4 pipelines rather than naive module partition on 2 modules will cause a speedup of 1.6x. Moreover, we can further accelerate the training loop by increasing the pipeline number (micro-batch number) to 8 and reaching a speedup of 1.77x. However, keep in mind that increasing pipeline number will increase memory usage due to versioned tensor.\par

@@

 However, we find out that 2 modules on Stacked Pipeline parallelism with each module running at Data Parallelism on 2 GPUs (Mixed Parallelism) still perform a little worse than Data parallelism on 4 GPUs. This result is anticipated because we perform our evaluation on a single GPU cluster with very high-speed inter-GPU connections, which makes the frequent tensor communication (ALLReduce) overhead introduced by Data Parallelism negligible. However, the cross-module dependencies for micro-batch will still cause some bubble time. As we mentioned before, the ideal network topologies for Pipeline parallelism are an extensive training system consisting of multiple GPU clusters connected through the network. Then we can use Data parallelism inside each cluster and Pipeline parallelism across clusters. We currently could not perform evaluation under this ideal scenario due to limited time and hardware resources, but we could predict that Pipeline parallelism mixed with Data parallelism will perform much better than only using Data parallelism in this scenario because Pipeline parallelism could reduce the communication overhead because inter-cluster communication only happens at module boundaries.\par
 The evaluation shows that our framework is flexible, and the user could switch to different combinations of Parallelism settings to reduce the total communication overhead and achieve optimal performance. It also demonstrates that our framework can unify both Spatial Partition and Pipeline Parallelism, validates the correctness of our training loop transformation, and offers a deeper insight into how to choose between different parallelism training settings.

>>>>

\end{figure} This overhead may be negligible for a GPU Cluster with high-bandwidth interconnections but will cause a significant performance drop for large distributed systems consisting of several nodes of GPU clusters with low bandwidth connection between them as shown in Figure \ref{fig:networktopo}. Because high bandwidth inter-GPU connection techniques like NV-Link can only scale to a limited number of devices, large training servers are often deployed across multiple nodes. It will be ideal to divide the model into several modules, assign each module to a cluster, use Spatial partition inside a GPU cluster, and adopt pipeline parallelism between modules.\par

@@

 AIRCoP \cite{aircopfeiwang} is a machine learning framework that supports all the parallel parallelisms above, therefore allows users to fully explore different parallel settings. AIRCoP implemented spatial partition before and our work implemented Pipeline parallelism for AIRCoP, thus make AIRCoP a unifying parallel training framework.\par

@@

 AIRCoP developed a formal syntax to represent tensor computation graphs. The syntax mainly has three components: Tensor and corresponding TensorType to represent input tensor and weight, Operation, and corresponding Operation Type to represent different Machine learning operations like Conv/MatMul/Relu and their input/output Type, Module to represent different modules. Operations are connected through let-bindings. Furthermore, we use annotation of the formal syntax to represent different parallelism strategies formally. We developed annotation systems for both Tensor, Operation, and Modules. Tensor and Operation annotation represents Spatial Partition by indicating the dimension to partition. Module annotation indicates how many micro-batches the mini-batch is split into. Users only need to write the sequential forward computation of the Machine Learning model, and the correct final training loop will be automatically computed through a series of transformations.

@@

 Simply splitting modules will cause large amount of stall time (or bubble time) due to complex dependencies between the sequential modules. As shown in Figure \ref{fig:multiplemodule}, if we naively spilt the model into 4 modules $m_0,m_1,m_2,m_3$. There will be sequential dependencies between the forward pass $F_0, F_1, F_2, F_3$ of modules $m_0,m_1,m_2,m_3$, and reverse dependencies between backward pass $B_3, B_2, B_1, B_0$. The dependencies between the forward pass and backward pass will force strict ordering between modules, introduce synchronizations at module boundaries, and cause significant idle time on 4 devices. \par

@@

 LMS (Lightweight Modular Staging) is a library-based staging compiler that distinguishes stages solely based on Type information. LMS is beneficial for developing DSL through overloading. It has a typed frontend making the code generation safe. LMS offers lifted (next-stage) versions of regular operators and standard control-flow statements, making symbolic code generation easy and convenient. AIRCoP is simply a tensor computation DSL embedded inside LMS framework.\par

@@

The Collection Phase collects the forward operations, the weight nodes, and the corresponding backward operation of each node. The backward operation is collected in reverse order to maintain the correct semantic of differentiation. Because the backward pass should be generated after the forward pass, we collect backward operation in the form of closures. The Collection Phase will also collect paired send/recv operations at the boundaries of different modules to maintain the correct data dependencies. We implement the Collection Phase by defining a function aircopCollect for each IR Node to collect the corresponding information. After the Collection Phase, we have the necessary information to perform automatic differentiation (AD). We also need to collect lifted tensors to generate the correct training loop. As we mentioned before, the forward pass and backward pass of micro-batches are iterated in different loops to achieve pipeline parallelism. Therefore, the tensors allocated in the forward pass and used in the backward pass will cause an error because forward and backward passes are now in different scopes. These tensors are called lifted tensors. We have to lift them into the outer scope and maintain multiple versions of them as an array of tensors. Micro-batches' forward and backward passes will use different versions of the lifted tensors. After collecting the lifted tensors, the Collection Phase is finished.\par

@@

The Generation Phase generates the correct training loop based on the collected information. Inside each module, it will first generate the top-level training loop for mini-batches. Then it will generate the forward training loop and backward training loop for micro-batches based on Module Annotations and use the collected backward operations to realize automatic differentiation as we mentioned in Section \ref{formalpipe}. Inside both the forward and backward pass, the read and write to the lifted tensors will be transformed into read/write to the versioned tensor array. The weight gradient will be accumulated in the backward pass and optimized at the end of the entire mini-batch.\par

@@

 This section shows the evaluation benchmarks of various forms of parallel training by running our generated CUDA code on 4 GeForce TITAN XP GPUs on server cuda.cs.purdue.edu. The server is a single GPU cluster with a high-speed inter-GPU connection. The ideal network topology for Pipeline parallelism is an extensive training system consisting of several Nodes, but we currently could not run the evaluation under that scenario due to limited hardware resources and time. The purpose of this benchmark is to validate our graph transformation and gain a deeper insight into the choice of different parallelism strategies.\par

@@

 The running time of 5 settings is shown in Figure \ref{fig:runtime}, the corresponding speedup is shown in Figure \ref{fig:speedup}. We can notice that simply dividing a model into 2 modules will not cause any speed up due to the low device usage caused by the cross-module dependencies. The device will idle while waiting for data to arrive. We can notice that using Pipeline parallelism with 4 pipelines rather than naive module partition on 2 modules will cause a speedup of 1.6x. Moreover, we can further accelerate the training loop by increasing the pipeline number (micro-batch number) to 8 and reaching a speedup of 1.77x. However, keep in mind that increasing pipeline number will increase memory usage due to versioned tensor.\par

@@

 However, we find out that 2 modules on Stacked Pipeline parallelism with each module running at Data Parallelism on 2 GPUs (Mixed Parallelism) still perform a little worse than Data parallelism on 4 GPUs. This result is anticipated because we perform our evaluation on a single GPU cluster with very high-speed inter-GPU connections, which makes the frequent tensor communication (ALLReduce) overhead introduced by Data Parallelism negligible. However, the cross-module dependencies for micro-batch will still cause some bubble time. As we mentioned before, the ideal network topologies for Pipeline parallelism are an extensive training system consisting of multiple GPU clusters connected through the network. Then we can use Data parallelism inside each cluster and Pipeline parallelism between clusters. We currently could not perform evaluation under this ideal scenario due to limited time and hardware resources, but we could predict that Pipeline parallelism mixed with Data parallelism will perform much better than only using Data parallelism in this scenario because Pipeline parallelism could reduce the communication overhead because inter-cluster communication only happens at module boundaries.\par
 The evaluation shows that our framework is flexible, and the user could switch to different combinations of Parallelism settings to reduce the total communication overhead and achieve optimal performance. It also demonstrates that our framework can unify both Spatial Partition and Pipeline Parallelism, validates the correctness of our training loop transformation, and offers a deeper insight into how to choose between different parallel training settings.
